---
output: html_document
---


```{r}
library(tidyverse)
library(dplyr)
library(readxl)
library(stringr)

county = read_csv('./data/countypres_2000-2016.csv')
unemployment = read_excel('./data/Unemployment.xls')
unemployment = unemployment %>% slice(7:n())
```

```{r}
colnames(unemployment) = unemployment[1,]
unemployment = unemployment[-1,]
```

#Task One: Done in Python#

**(i) proportion of the vote won by either Democartic or Republican candidate**

### Import the required packages and dataset###

```{python}
import pandas as pd
county = pd.read_csv('./data/countypres_2000-2016.csv')
unemployment = pd.read_excel('./data/Unemployment.xls')
```

##**to get the elections percent firstly**##

### select the rows in unemployment dataset###

```{python}
unemployment = unemployment[6:]
unemployment.columns = unemployment.iloc[0]
unemployment = unemployment.drop(unemployment.index[0])
unemployment.reset_index(inplace=True)
```

### use groupby to get the vote percent in each year for each party###

```{python}

vote_percent_df = county.groupby(['state_po','county','party','year']).apply(lambda g: g.candidatevotes/g.totalvotes)
vote_percent_df = vote_percent_df.reset_index().drop('level_4',axis=1)
vote_percent_df.rename(columns={ vote_percent_df.columns[-1]: "vote_percent" }, inplace = True)
```

### filter the vote percent with year >= 2018 and party type = democrat###
```{python}
vote_percent_df = vote_percent_df.loc[(vote_percent_df['year'] >= 2008) & (vote_percent_df['party'] == 'democrat')]
vote_percent_df.reset_index(drop=True,inplace=True)
```

##**Then to get the unemployment rates**##
### There are several rows needed to be moved like area_name = 'United States' etc###

```{python}
unemployment = unemployment[unemployment['Rural_urban_continuum_code_2013'].notna()]
selected_column = ['area_name','Unemployment_rate_2008','Unemployment_rate_2012','Unemployment_rate_2016']
selected_unemployment = unemployment[selected_column].reset_index(drop=True)
```

### split area_name into state and county###

```{python}
selected_unemployment[['State','County']] = selected_unemployment['area_name'].str.split(',',expand=True)
selected_unemployment.drop('area_name',inplace=True,axis=1)
selected_unemployment
```

### rename column to be precisely###

```{python}
selected_unemployment.columns = ['2008','2012','2016','State','County']
selected_unemployment
```

#Task Two: Done in R:

##**(i) Proportion of vote won**

```{r}
selected_county = county %>% group_by(year,state_po,party,county) %>% mutate(percent = candidatevotes/totalvotes) %>% select(year,state_po,FIPS,county,party,percent)
selected_county = na.omit(selected_county[selected_county$party == 'democrat',])
selected_county$FIPS = str_pad(selected_county$FIPS,max(nchar(selected_county$FIPS)),side="left", pad="0")
selected_county = selected_county %>% filter(year >= 2008)
```

```{r}
selected_column = c('FIPStxt','area_name','Unemployment_rate_2008','Unemployment_rate_2012','Unemployment_rate_2016')
selected_unemployment = unemployment[selected_column]
selected_unemployment = na.omit(selected_unemployment %>% separate(area_name,c('County','State'),','))
```

```{r}
library(sf)
library(tmap)
library(tmaptools)
```

```{r}
usgeo <- read_sf('./data/tl_2019_us_county-1/tl_2019_us_county.shp')

```

### to draw the Texas firstly (Texas FIPS = 48)###
### draw the shape firstly###

```{r}
texas_geo <- filter(usgeo,STATEFP=='48')
qtm(texas_geo)

```
### Then we can draw the percent of vote in texas###

```{r}
texas_geo$NAME <- as.character(texas_geo$NAME)
texas_map <- left_join(texas_geo, selected_county,by=c("GEOID"="FIPS"))
tx_plot = tm_shape(texas_map) +
tm_fill("percent", title="percent of votes for democrat in Texas", palette = "PRGn") +
tm_borders(alpha=.5) +
  tm_style('classic') +
  tm_facets(by=c('year'),ncol=1) +
  tm_layout(main.title = 'Percent of votes for democrat in Texas',main.title.size=1,legend.outside = TRUE)
tx_plot
```


### to draw the Calidornia secondly (California FIPS = 06)###
### also draw the shape firstly###

```{r}
ca_geo <- filter(usgeo,STATEFP=='06')
qtm(ca_geo)
```

### Then draw the percent of vote in CA###

```{r}
ca_geo$NAME <- as.character(ca_geo$NAME)
ca_map <- left_join(ca_geo, selected_county,by=c("GEOID"="FIPS"))
ca_plot = tm_shape(ca_map) +
tm_fill("percent", title="percent of votes for democrat in Calidornia", palette = "PRGn") +
tm_borders(alpha=.5) +
  tm_facets(by=c('year'),ncol=1) +
  tm_style('classic') +
  tm_layout(main.title = 'Percent of votes for democrat in California',main.title.size=1,legend.outside = TRUE)
ca_plot
```

### draw Pennsylvania lastly(FIPS = 42)###
### to draw the shape of Penn###

```{r}
penn_geo <- filter(usgeo,STATEFP=='42')
qtm(penn_geo)
```

### Then draw the percent of vote in Penn###

```{r}
penn_geo$NAME <- as.character(penn_geo$NAME)
penn_map <- left_join(penn_geo, selected_county,by=c("GEOID"="FIPS"))
penn_plot = tm_shape(penn_map) +
tm_fill("percent", title="percent of votes for democrat in Pennsylvania", palette = "PRGn") +
  tm_facets(by=c('year'),ncol=1) +
tm_borders(alpha=.5) +
  tm_style('classic') +
  tm_layout(main.title = 'Percent of votes for democrat in Pennsylvania',main.title.size=1,legend.outside = TRUE)
penn_plot
```

##**Task 2: draw for Unemployment Rate**##

### firstly, to let unemployment rate dataset to be tidy dataset###

```{r}
library(tidyr)

selected_unemployment = pivot_longer(selected_unemployment, cols=4:6, names_to = "year", values_to = "Unemployment Rate")

selected_unemployment$year = substr(selected_unemployment$year,19,23)

selected_unemployment$year = as.numeric(selected_unemployment$year)
```

### unemployment rate in Texas###

```{r}
texas_map <- left_join(texas_geo, selected_unemployment,by=c("GEOID"="FIPStxt"))
texas_map$`Unemployment Rate` <- as.numeric(texas_map$`Unemployment Rate`)
tx_un_plot = tm_shape(texas_map) +
tm_fill('Unemployment Rate', title="Unemployment Rate in Texas(%)", palette = "PRGn") +
tm_borders(alpha=.5) +
  tm_style('classic') +
  tm_facets(by=c('year'),ncol=1) +
  tm_layout(main.title = 'Unemployment rate in Texas',main.title.size=1,legend.outside = TRUE)
tx_un_plot
```

### Unemployment rate in CA###

```{r}
ca_map <- left_join(ca_geo, selected_unemployment,by=c("GEOID"="FIPStxt"))
ca_map$`Unemployment Rate` <- as.numeric(ca_map$`Unemployment Rate`)
ca_un_plot = tm_shape(ca_map) +
tm_fill('Unemployment Rate', title="Unemployment Rate in California(%)", palette = "PRGn") +
tm_borders(alpha=.5) +
  tm_style('classic') +
  tm_facets(by=c('year'),ncol=1) +
  tm_layout(main.title = 'Unemployment rate in California',main.title.size=1,legend.outside = TRUE)
ca_un_plot
```

### Unemployment in Penn###

```{r}
penn_geo$NAME <- as.character(penn_geo$NAME)
penn_map <- left_join(penn_geo, selected_unemployment,by=c("GEOID"="FIPStxt"))
penn_map$`Unemployment Rate` <- as.numeric(penn_map$`Unemployment Rate`)
penn_un_plot = tm_shape(penn_map) +
tm_fill("Unemployment Rate", title="Unemployment Rate in Pennsylvania(%)", palette = "PRGn") +
  tm_facets(by=c('year'),ncol=1) +
tm_borders(alpha=.5) +
  tm_style('classic') +
  tm_layout(main.title = 'Unemployment Rate in Pennsylvania',main.title.size=1,legend.outside = TRUE)
penn_un_plot
```

## Finally, combine plots###

### For Pennsylvania###

```{r}
current.mode <- tmap_mode('plot')
tmap_arrange(penn_plot,penn_un_plot,widths = c(1,1))
tmap_mode(current.mode)
```

### For Texas###

```{r}
current.mode <- tmap_mode('plot')
tmap_arrange(tx_plot,tx_un_plot,widths = c(1,1))
tmap_mode(current.mode)
```

### For California###

```{r}
current.mode <- tmap_mode('plot')
tmap_arrange(ca_plot,ca_un_plot,widths = c(1,1))
tmap_mode(current.mode)
```