---
title: "ANLY503_Q1"
output: html_document
---

```{r setup, include=FALSE}

# load the required packages
library(dplyr)
library(ggplot2)
library(visdat)
library(hrbrthemes)
library(naniar)
library(Hmisc)
library(skimr)
```


# Question One:


## <span style="color:red">R Version:</span>

we need to load the file firstly

```{r}
# load the dataset
transactions = read.csv('./data/transactions.csv')
accounts = read.csv('./data/accounts_analytical.csv')
```

and then take a glance for these datasets

```{r}
# to find descriptive information for transactions
Hmisc::describe(transactions)
```

```{r}
glimpse(accounts)
```

 because there are too much variables with long name

 I make a subset dataset for columns that have null values

```{r}
# do analysis for null values
accounts_with_null_columns = accounts %>% select_if(funs(sum(as.numeric(is.na(.)), na.rm=TRUE) != 0))

theme_set(theme_ipsum())
vis_dat(accounts_with_null_columns,warn_large_data=FALSE)
```

 we can find there are lots of NULL values in columns related with loan

```{r}
gg_miss_var(transactions)
```

there are only one column have null values that is 'account'

## Question 1A: Whether an account has a credit card or not

```{r}

# filter the dataset to find the accounts with or without credit cards
accounts$credit_cards[is.na(accounts$credit_cards)] <- 0

p <- accounts %>%
  ggplot( aes(x=credit_cards)) +
    geom_histogram( binwidth=1, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
    ggtitle("The count plot of credit cards") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=15)
    )

p
```

there are 3775 accounts do not have a credit card

## Question 1B: Whether an account has a loan or not

```{r}
# find whether account has loan or not 
accounts$loan_status[accounts$loan_status == ''] <- 'Unknown'

ggplot(accounts,aes(loan_status,fill=loan_status)) +
  geom_bar() +
  labs(x=NULL,y='Count') +
  geom_text(stat='count',aes(label = after_stat(count),vjust=-1,size=3.5)) +
  ggtitle('The count plot of loan status')
```

we can there are lots of unknow values, and there are 448 loan is active, and 234 loan is expired

## Question 1C: The average balance for the account

```{r}

# calculate the average balance
accounts$mean_balance = (accounts$max_balance + accounts$min_balance) / 2

# to show the distribution of average balance for accounts
average_balance <- ggplot(accounts,aes(x=mean_balance)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 10000,
                  colour = 'black', fill = 'white') +
  geom_density(alpha=0.2,fill='blue') +
  ggtitle('The distribution of average balance')

average_balance
```

## Question 1D:

```{R}
average_balance + facet_wrap(~credit_cards+loan_status,nrow=4,
                             labeller = labeller(
                               credit_cards = c('0' = "0 Credit Card",
                                                '1' = "1 Credit Card",
                                                '2' = "2 Credit Cards")
                             )) +
  ggtitle('The distribution of average balance under different number of credit cards')
```

there are no obvious change of distribution of average balance by the number of credit cards

```{r}
# to make a heatmap to find the count under different credit_cards and loan_status
counts = count(accounts,loan_status,credit_cards) 

ggplot(counts,aes(x=loan_status,y=credit_cards)) +
  geom_tile(aes(fill=n)) +
  geom_text(aes(label=n),stat='identity') + 
  scale_fill_continuous(low="grey", high="pink", limits=c(0,4000))+
  ggtitle('number of accounts type')
```

after plotting this matrix, we can find most accounts do not have credit_cards and have Unknow loan status

## <span style="color:red">Python Version:</span>

```{r}
library(reticulate)
```

### import the required packages

```{python}
# loan packages
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
```

### load the datasets

```{python}
transactions = pd.read_csv('./data/transactions.csv')
accounts = pd.read_csv('./data/accounts_analytical.csv')
```

### Information for transactions

```{python}
# take some descriptive information
transactions.describe()

transactions.info()
```

### Infotmation for accounts

```{python}
accounts.describe()

accounts.info()
```

## Question 1A:
### Whether an account has a credit card or not

```{python}
# filter the dataset 

accounts['credit_cards'] = accounts['credit_cards'].fillna('Unknow')

sns.set_theme(style='darkgrid')

ax = sns.countplot(x="credit_cards", data=accounts)

ax.set_xticklabels(['One Credit Card','Two Credit Cards','Unknown'])
ax.set_title('The countplot of credit_cards')
ax.set_xlabel('Number of credit cards')
```

### most accounts do not have credit card information, and second highest is one credit card account

## Question 1B:
### Whether an account has a loan or not

```{python}
accounts.loan_status = accounts.loan_status.fillna('Unknown')

sns.set_theme(style='darkgrid')

ax = sns.countplot(x="loan_status", data=accounts)

ax.set_title('The countplot of loan_status')
```

same as the last plot, most accounts do not have loan status information, and second highest is current loan.

## Question 1C:
### The average balance for the account

```{python}
accounts['mean_balance'] = (accounts['max_balance'] + accounts['min_balance']) / 2

f, axes = plt.subplots(1, 2)

sns.boxplot(x='mean_balance',data=accounts,ax=axes[0])
axes[0].title.set_text('The Boxplot of Mean Balance')
sns.distplot(x=accounts.mean_balance,ax=axes[1])
axes[1].title.set_text('The KDE plot of Mean Balance')
axes[1].set_xticklabels(labels='mean balance')
plt.tight_layout()
plt.show()
```

### we can find most mean_balance is around $25000.

## Question 1D:

### to combine these three questions together

```{python}

# to make the boxplot and facetgrid under different number of credit_cards and loan_status
ax = sns.boxplot(x="loan_status", y="mean_balance", hue="credit_cards",
                 data=accounts, palette="Set3")
ax.legend(loc='center right', bbox_to_anchor=(1.25, 0.5), ncol=1)
ax.set_title('The Boxplot of mean balance in loan_status and number of credit cards')
```
