---
title: "ANLY503_Q4"
output: html_document
---


```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(visdat)
library(hrbrthemes)
library(naniar)
library(Hmisc)
library(skimr)
library(lubridate)
```

# Question Four:


## <span style="color:red">R Version:</span>

```{r}
transactions = read.csv('./data/transactions.csv')
accounts = read.csv('./data/accounts_analytical.csv')
```

### Find the account id has the highest transaction amount

```{r}
# filter the dataset to find the highest transaction amount
account_id = transactions %>% group_by(account_id) %>% mutate(cnt = n()) %>% ungroup %>% filter(cnt == max(cnt))
#account_id
```

##**Plot One: Smooth Line by year**

```{r}
# plot the amount by smooth line
account_id$date = ymd(account_id$date)
account_id %>% group_by('date') %>% 
ggplot(aes(x=date,y=amount,group=type,color=type))+
  geom_smooth()
  
```

##**Plot Two:Different Type of transaction in different year**

```{r}
# to plot the time series of transaction in credit and debit
account_id['year'] = year(account_id$date)
account_id['month'] = month(account_id$date)
account_id['month'] = factor(account_id$month, levels = unique(account_id$month))
ggplot(account_id,aes(x=month,y=amount,group=type,color=type)) +
  facet_grid(year~.) +
  geom_line() + 
  ggtitle('The Time Series Plot of Type of Transactions in each year')
```

## **Plot Three: Time Series by Method**

```{r}
account_id$method[grepl("^ *$",account_id$method)] <- NA
ggplot(account_id,aes(x=month,y=amount,group=method,color=method)) +
  facet_grid(year~.,space = "free") +
  geom_line(na.rm=TRUE) + 
  ggtitle('The Time Series Plot of Method of Transactions in each year')
```
  
## **Plot Four: The Time Series Plot of Category**  
  
```{r}
account_id$category[account_id$category == "" ] <- 'Unknown'
ggplot(account_id,aes(x=month,y=amount,group=category,color=category)) +
  facet_grid(year~.,space = "free") +
  geom_line() + 
  theme(text = element_text(size = 8), element_line(size = 0.6))
  ggtitle('The Time Series Plot of Method of Transactions in each year')
```

## Conclusion:

For the last plot, we can find the household payment, insurance payment, loan payment, intest credit, and statement charge all have frequency, and the transaction amount is constant.

for the first and the second plot, when time is close to the end of month, the amount will increase.


## <span style="color:red">Python Version:</span>

```{python}
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import datetime as dt
```

### load the datasets

```{python}
transactions = pd.read_csv('./data/transactions.csv')
accounts = pd.read_csv('./data/accounts_analytical.csv')
```

```{python}
# group the dataset by account_id
account_id = transactions.groupby('account_id').count().reset_index()[['account_id','id']].sort_values(ascending=False,by='id')['account_id'].reset_index()
account_df = transactions[transactions['account_id'] == account_id['account_id'][0]]

#account_df
```

## **Plot One: Amount versus Type**

```{python}
# find the trend in each year
# parse the date
account_df['year'] = pd.to_datetime(account_df['date']).dt.year
account_df['month'] = pd.to_datetime(account_df['date']).dt.month

f,ax = plt.subplots(2,1,figsize=(12,12))
sns.lineplot(x='month',y='amount',hue='year',data=account_df[account_df['type']=='credit'],ci=None,ax=ax[0],palette='Set1')
ax[0].set_title('The trend of amount from 1993 to 1998 for credit')

sns.lineplot(x='month',y='amount',hue='year',data=account_df[account_df['type']=='debit'],ci=None,ax=ax[1],palette='Set1')
ax[1].set_title('The trend of amount from 1993 to 1998 for debit')
```

## **Plot Two: Amount versus Method**

```{python}
f,ax = plt.subplots(2,1,figsize=(12,10))
sns.lineplot(x='month',y='amount',hue='year',data=account_df[account_df['method']=='bank transfer'],ci=None,ax=ax[0],palette='Set1')
ax[0].set_title('The trend of amount from 1993 to 1998 for Bank Transfer')

sns.lineplot(x='month',y='amount',hue='year',data=account_df[account_df['method']=='cash'],ci=None,ax=ax[1],palette='Set1')
ax[1].set_title('The trend of amount from 1993 to 1998 for Cash')

```

## **Plot Three: Amount versus Category**

```{python}
f,ax = plt.subplots(12,5)
account_df['category'].fillna('Unknown',inplace=True)
sns.relplot(x='month',y='amount',col='year',hue='category',data=account_df,size='category',ci=None,kind='line')
plt.title('The Transaction amount versus Category')
```

## Conclusion:

The Conclusion is the same as R version.