---
title: "ANLY503_Q2"
output: html_document
---

```{r setup, include=FALSE}
# loan the packages
library(dplyr)
library(ggplot2)
library(visdat)
library(hrbrthemes)
library(naniar)
library(Hmisc)
library(skimr)

theme_set(theme_classic()+theme(axis.text = element_text(size=14),
                                axis.title = element_text(size=16),
                                legend.text = element_text(size=14),
                                legend.title = element_text(size=16),
                                plot.title = element_text(size=18),
                                plot.subtitle = element_text(size=16),
                                plot.caption = element_text(size=12)))
```

# Question Two:


## <span style="color:red">R Version:</span>

```{r}
# load datasets
transactions = read.csv('./data/transactions.csv')
accounts = read.csv('./data/accounts_analytical.csv')
```

### What is the distribution of all loans and what are their characteristics?

the distribution of loan_amount

**Plot 1: loan_amount versus loan_status**

```{r}
# to plot the distribution of loan_amount under different loan_status
ggplot(accounts, aes(x=loan_amount, fill=loan_status)) +
  geom_histogram(alpha=.5, position="identity") +
  ggtitle('plot of distribution of loan_amount and loan_status')
```

**Plot 2: distribution of loan_term**

```{r}
# count plot of loan_term
loan_term_counts = as.data.frame(table(accounts$loan_term))
ggplot(loan_term_counts,aes(x=as.factor(Var1),y=as.factor(Freq),fill=Freq)) +
  geom_bar(stat='identity') +
  scale_color_brewer(palette="Dark2") +
  xlab("loan_term") +
  ylab("count") +
  ggtitle('the frequency of loan term')
```

**Plot 3: Boxplot of loan_default, and loan_amount**

```{r}
# to plot the boxplot of loan_default and loan_amount
accounts$loan_default[is.na(accounts$loan_default)] <- 'Unknown'

bp<-ggplot(accounts, aes(x=loan_default, y=loan_amount, fill=loan_default)) +
  geom_boxplot()
bp + scale_fill_brewer(palette="Dark2") +
  ggtitle('Boxplot of loan amount under different loan_default')
```

### Conclusion:
1. the distribution of loan_amount under different loan_status is similar

   but loan_status == True will have kind of higher loan_amount.

2. we can find loan_term = 36 months has the least count,

   but other loan_term has simialr counts by the distribution of loan_term 

3. we can find if loan_default = True, account will have higher loan_amount

   than account does not have default loan
   
   
## <span style="color:red">Python Version:</span>

```{python}
# import packages
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
```

### load the datasets

```{python}
transactions = pd.read_csv('./data/transactions.csv')
accounts = pd.read_csv('./data/accounts_analytical.csv')
```

**Plot One: Loan Term versus Loan Status**

```{python}
# use loan_status as hue to plot loan term
ax = sns.countplot(x='loan_term',hue='loan_status',data=accounts,palette='Set2')
ax.legend(loc='center right', bbox_to_anchor=(1.15, 0.5), ncol=1)
ax.set_title('the loan term under different loan status')
```

**Plot Two: Loan Default versus Loan Amount**

```{python}
ax = sns.boxplot(x='loan_default',y='loan_amount',data=accounts,palette='Set1')
ax.set_title('Loan Amount under default loan account and not default loan account')
ax
```

### Conclusion:

1. most loan_status = expired account has short loan term like 12 and 24 months, 

and most loan_status = current has long loan term like 48 and 60 months

2.loan_default = True account have higher loan amount than loan_default = False account.