---
title: "ANLY503_Q3"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(visdat)
library(hrbrthemes)
library(naniar)
library(Hmisc)
library(skimr)
```

# Question Three:


## <span style="color:red">R Version:</span>

```{r}
transactions = read.csv('./data/transactions.csv')
accounts = read.csv('./data/accounts_analytical.csv')
```

## **Plot One:Account Open Date versus Loan**

Parse the date firstly

```{r}
# loan packages to parse the datetime firstly
library(lubridate)
accounts['accounts_year'] = year(accounts$acct_creation_date)
accounts['accounts_day'] = day(accounts$acct_creation_date)
accounts['accounts_weekend'] <- accounts$day %in% c("Sat", "Sun")
accounts['month'] <- month(accounts$acct_creation_date)
accounts['accounts_holiday'] <- with(accounts,ifelse(
                                    month >= 2 & month <= 5, 'Spring', ifelse(
                                    month >= 6 & month <= 8, 'Summer', ifelse(
                                    month >= 9 & month <= 11,'Autumn', 'Winter')
                                    )
                                    )
)

# filter the dataset
accounts$loan_date[grepl("^ *$",accounts$loan_date)] <- NA

accounts['loan_year'] = year(accounts$loan_date)
accounts['loan_day'] = day(accounts$loan_date)
accounts['loan_weekend'] <- accounts$day %in% c("Sat", "Sun")
accounts['month'] <- month(accounts$loan_date)
accounts['loan_holiday'] <- with(accounts,ifelse(
                                    month >= 2 & month <= 5, 'Spring', ifelse(
                                    month >= 6 & month <= 8, 'Summer', ifelse(
                                    month >= 9 & month <= 11,'Autumn', 'Winter')
                                    )
                                    )
)
```

```{r}
# to count the accounts in each year
library("viridis")
ggplot(data=accounts,aes(accounts_year)) +
  geom_bar(stat='count',color='tan3',fill='slateblue1') +
  ggtitle('The count of accounts created in each year')
```


```{r}
# to plot two barcharts of loan_status in each year
par(mfrow=c(1,2))
year_loan_status_counts = accounts %>% group_by(accounts_year,loan_status) %>% count()
year_loan_status_counts$loan_status = sub("^$", "Unknown", year_loan_status_counts$loan_status)
ggplot() +
    geom_bar(data=year_loan_status_counts,aes(fill=loan_status,x=accounts_year,y=n),stat='identity',
             position = position_dodge(width = 0.8), binwidth = 25) +
    labs(x = "Year") +
    ggtitle('The Loan Status under different year')

year_loan_default_counts = accounts %>% group_by(accounts_year,loan_default) %>% count()
year_loan_default_counts$loan_default[is.na(year_loan_default_counts$loan_default)] <- 'Unknown'
ggplot() +
    geom_bar(data=year_loan_default_counts,aes(fill=loan_default,x=accounts_year,y=n),stat='identity',
             position = position_dodge(width = 0.8), binwidth = 25) +
    labs(x = "Year") +
    ggtitle('The Loan Default in different seasons')

```

## **Plot Two:Loan Open Date versus  Loan**

```{r}
# to plot two barcharts of loan status by loan open date
par(mfrow=c(1,2))
year_loan_status_counts = accounts %>% group_by(loan_year,loan_status) %>% count()
year_loan_status_counts$loan_status = sub("^$", "Unknown", year_loan_status_counts$loan_status)
ggplot() +
    geom_bar(data=year_loan_status_counts,aes(fill=loan_status,x=loan_year,y=n),stat='identity',
             position = position_dodge(width = 0.8), binwidth = 25) +
    scale_y_continuous(limits=c(0,800)) +
    labs(x = "Year") +
    ggtitle('The Loan Status under different year')

year_loan_default_counts = accounts %>% group_by(loan_year,loan_default) %>% count()
year_loan_default_counts$loan_default[is.na(year_loan_default_counts$loan_default)] <- 'Unknown'
ggplot() +
    geom_bar(data=year_loan_default_counts,aes(fill=loan_default,x=loan_year,y=n),stat='identity',
             position = position_dodge(width = 0.8), binwidth = 25) +
    scale_y_continuous(limits=c(0,800)) +
    labs(x = "Year") +
    ggtitle('The Loan Default in different year')
```

## Conclusion:

1. The status and default of loan is not related with the account created date.

2. The status of loan is related with loan created date, because when the time passed, the loan will be more possible not to be default, expecially for current loan.
   
  
## **Plot Three: specific set of accounts that seem to be at higher or lower risk of defaulting**

to begin with, let me find whether there is any similarity within default loan accounts 

```{r}
#accounts %>% filter(loan_default=='TRUE') %>% head(5)
```
   
The accounts that have default loan all have negative min_balance.

## The accounts have higher risk of default loan are:

```{r}
#accounts %>% filter(min_balance <= 0)
```
```{r}
# to plot the pie chart of percentage of higher risk accounts
library(scales)
higher_risk = accounts %>% filter(min_balance <= 0)
higher_risk$loan_default[is.na(higher_risk$loan_default)] <- 'Unknown'

higher_risk = higher_risk %>%
  group_by(loan_default) %>%
  summarise(counts = n()) %>%
  mutate(percentages = round(counts/sum(counts)*100,2))

ggplot(data=higher_risk,aes(x=factor(1),y=counts,fill=loan_default)) +
  geom_col(aes(fill = loan_default), width = 1) +
  geom_text(aes(label = percentages), position = position_stack(vjust = 0.5)) +
  coord_polar("y") + ggtitle('The distribution of higher default loan risk accounts')
```

## Conclusion:

So in the pie chart above, the accounts all are higher risk default loan.

## **Plot Four: The accounts have higher risk of default loan are**

Follow the same rule, we try to find the accounts with Top 10% min_balance

```{r}
n = 15

Top_10_min_balance = subset(accounts, min_balance > quantile(min_balance, prob = 1 - n/100))

ggplot(data=Top_10_min_balance,aes(loan_default)) +
  geom_bar(stat='count')
```

None of them have default loan.

## Conclusion:

I think min_balance is importance factor to determine whether account is under higher risk of default loan.


## <span style="color:red">Python Version:</span>

```{python}
# loan packages
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import datetime as dt
```

### load the datasets

```{python}
transactions = pd.read_csv('./data/transactions.csv')
accounts = pd.read_csv('./data/accounts_analytical.csv')
```

## **Plot One: Account Open Date versus Loan**

Parse the date firstly

```{python}
accounts['accounts_year'] = pd.to_datetime(accounts.acct_creation_date).dt.year
accounts['loan_year'] = pd.to_datetime(accounts.loan_date).dt.year
```

```{python}
# to plot countplot of loan_default under each year
ax = sns.countplot(x='accounts_year',hue='loan_default',data=accounts)
ax.set_title('The count plot of loan_default from 1993 to 1997')
```

```{python}
ax = sns.catplot(x = 'accounts_year',hue='loan_default',col='loan_status',
                data = accounts, kind = 'count',
                height = 6, aspect = .7,palette='Set2')
plt.show()
  
```

## Conclusion:

For loan_status = expired account, the loan default is being less when the accounts_year higher.

## **Plot Two: Loan Date versus Loan Default**

```{python}
# plot barchart for loan_default
ax = sns.catplot(x = 'loan_year',hue='loan_default',col='loan_status',
                data = accounts, kind = 'count',
                height = 6, aspect = .7,palette='Set1')
plt.show()
```

## Conclusion:

The loan default is more related with the loan created date.

especially for loan_status = current, when time passes by, account will be less likely to be default loan.

## **Plot Three: Specific account has higher or lower risk of default loan**

### Lower Risk

```{python}
lower_risk = accounts.nlargest(100,'min_balance')['loan_default']
lower_risk.fillna('Unknown',inplace=True)
lower_risk.value_counts().plot.pie(autopct='%1.0f%%')
plt.title('Pie Chart for the count of loan_default in accounts that have top 100 highest min_balance')
```

### Higher Risk:

```{python}
# find the percentage of higher risl
higher_risk = accounts[accounts['min_balance'] < 0]
higher_risk['loan_default'].fillna('Unknown',inplace=True)
higher_risk['loan_default'].value_counts().plot.pie(colors=['c','b'],shadow=True,explode=(0.1,0),autopct='%1.0f%%')
```
## Conclusion:

same as R version, I think min_balance is importance factor to determine whether account is under higher risk of default loan. Higher min_balance will have lower default loan risk.

